---
title: "04 Mini Project"
author: Solveig Senf
format:
  html: default
editor_options: 
  chunk_output_type: console
---

# Introduction

This project uses data on the beloved 2000s sitcom (and my personal favorite TV show), Gilmore Girls. This data set, found on Kaggle and accessed [here](https://www.kaggle.com/datasets/julqka/gilmore-girls-lines?resource=download), contains the full script from all seven seasons of the show. 

First, I clean the data, making a few different data frames for different text analysis purposes. To do so, I make use of multiple str_ functions: str_to_lower, str_view, str_replace_all, and str_to_title. 

Then, I use the bing sentiment application to analyze the positivity of the show as a whole. After removing the character names and stop words, I found the 10 most used positive words and the 10 most used negative words. Interestingly, positive words were used much more frequently than negative words, providing insight into the attitude of the show. 

Next, I used a similar application, the afinn application,  to analyze the positivity of the five main characters in the show for each season. Generally, Luke tends to be the most negative character, whereas Sookie tends to be one of the most positive characters. From this analysis, we can also see how characters shift in positivity levels from season to season. 
I also used the data to make a word cloud of the 100 most commonly used words in the show. The word cloud demonstrates the casual language used by the characters and displays some Gilmore Girl classics like “coffee”, “mother”, and “Yale”. 

Finally, I wanted to run a text analysis to help further the contentious, and VERY serious, debate on which of Rory’s boyfriends was the best. First, I used the str_count function to count how many times Rory mentioned each of their names, perhaps indicating the length and importance of each relationship. Then, I plotted the mean positivity score (using the afinn application) for each boyfriend, providing evidence for which of them was the best fit. Was Dean the nicest? Did Logan treat Rory right? Was Jess all bad, or did he have a kind side? This plot will help answer some of these persisting questions. 


```{r}
#| include: FALSE

library(tidyverse)
library(tidytext)
library(textdata)
library(wordcloud)
library(wordcloud2)
library(viridis)
library(ggthemes)
library(gutenbergr)
library(readr)
library(ggplot2)
library(wesanderson)

gilmore_girls <- read_csv("/Users/bethsenf/SDS264/Data/Gilmore_Girls_Lines.csv")
```

# Data Cleaning

```{r}
gilmore_girls_clean <- gilmore_girls |>
  mutate(line = str_to_lower(Line), #first str_function
         character = str_to_lower(Character), 
         character = str_replace_all(character, "/", " ")) |> #second str_function, first regular expression ("/")
  select(-Character, -Line, -`...1`) |>
  rename(season = Season) |>
  unnest_tokens(word, line, token = "words") #makes each word its own row
```


str_view(gilmore_names$character, "^.*/.*$") #ensure the str_replace_all worked, third str_function, second regular expression ("^.*/.*$") (Removed from R chunch to render.)


```{r}
gilmore_names <- gilmore_girls_clean |>
  select(-word, -season) |>
  distinct(character) #makes a list of all character names from the script (includes extras with lines)

gilmore_no_names <- gilmore_girls_clean |>
  anti_join(gilmore_names, by = c("word" = "character")) 
```

```{r}
gilmore_main_characters <- gilmore_girls_clean |>
  count(character) |>
  arrange(desc(n)) |>
  slice_max(n, n = 5) #makes a list of the main characters, determined by number of words spoken
```

```{r}
smart_stopwords <- get_stopwords(source = "smart")

gilmore_no_stop_names <- gilmore_no_names|>
  anti_join(smart_stopwords) 
```

# Positivity Analysis

## The Show

Is Gilmore Girls truely a feel-good TV show, or does it have a dark side?

```{r}
bing_sentiments <- get_sentiments(lexicon = "bing") #indicates positive/negative for each word

bing_gilmore <- gilmore_no_stop_names |>
  inner_join(bing_sentiments) |>
  count(sentiment, word, sort = TRUE) 

bing_gilmore |>
  group_by(sentiment) |>
  summarize(count = sum(n))

bing_gilmore|>
  group_by(sentiment) |>
  slice_max(n, n = 10) |> #10 most used positive and negative words
  ungroup() |>
  ggplot(aes(x = fct_reorder(word, n), y = n, fill = sentiment)) +
    geom_col() +  
    coord_flip() +
    facet_wrap(~ sentiment, scales = "free") +
    labs(x = "Count",
         y = "Word") +
    scale_fill_brewer(palette = "YlOrRd", direction = -1) +
    theme_classic()
```

## The Characters

Is Emily the villain after all? Looks like Luke was the true downer!

```{r}
afinn_sentiments <- get_sentiments(lexicon = "afinn") #gives each word a positivity score 

gilmore_no_stop_names |>
  inner_join(afinn_sentiments, relationship = "many-to-many") |>
  semi_join(gilmore_main_characters) |> #takes only words from the main characters
  mutate(character = str_to_title(character)) |> #fourth str_function
  group_by(character, season) |>
  summarize(mean_value = mean(value)) |>
  ungroup() |>
  ggplot(aes(x = character, y = mean_value, fill = character)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~season) +
    coord_flip() +
    scale_fill_brewer(palette = "OrRd") + #Fall color palette
    theme_classic() +
    labs(title = "Positivity of Main Characters by Season",
       y = "Positivity Score",
       x = "Character")
```

# Word Cloud

Gilmore Girls is known for its witty characters, but what words do they actually frequent? *Yeah,* it looks like this word cloud will help us figure this out!

```{r}
gilmore_words <- gilmore_no_stop_names |>
  count(word) |>
  arrange(desc(n)) |>
  anti_join(stop_words)

gilmore_words_df <- gilmore_words |>
  slice_head(n = 100) |>
  data.frame()

wordcloud2(
  gilmore_words_df, 
  size = 1, 
  shape = 'circle',
  minSize = .5,
  color = "random-dark"
)
```

# Boyfriend Analysis

## Boyfriend Mentions

Did Rory have a favorite the whole time? Was she more excited to talk about one boyfriend compared to the others? It looks like Dean and Logan are separating themselves in the competition for best boyfriend! 

```{r}
gilmore_girls_clean |>
  filter(character == "rory") |>
  summarize(jess_mentions = sum(str_count(word, "^jess")), #fifth str_function, thrid regular expression
            logan_mentions = sum(str_count(word, "^logan")),
            dean_mentions = sum(str_count(word, "^dean")))
```

## Positivity of Boyfriends

Rory might have mention one boyfriend more than the others, but what about how *they* treated *her*? Dean and Logan continue to be great contenders for the title of Rory's best boyfriend. Unfortunately, the results of this affin text analysis do not point to a decisive winner. Jess needs some defending, but it looks like the debate will continue. 

```{r}
gilmore_no_stop_names |>
  inner_join(afinn_sentiments, relationship = "many-to-many") |>
  filter(character == c("jess", "dean", "logan")) |> #takes only words from rory's boyfriends
  group_by(character) |>
  summarize(mean_value = mean(value)) |>
  ungroup() |>
  ggplot(aes(x = character, y = mean_value, fill = character)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = round(mean_value, 2)), vjust = -.5) +
    scale_fill_brewer(palette = "OrRd") +
    theme_classic() +
    labs(title = "Positivity of Rory's Boyfriends",
       y = "Positivity Score",
       x = "Character") 
```

